---
title: "Bobcat-Coyote Overlap"
author: "Viorel Popescu"
date: "2025-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("~/Downloads/6.co-occupancy/My Data")
```

```{r}
library(tidyverse)
library(dplyr)
library(lubridate)
library(chron)
library(camtrapR)
library(epitools)
library(pivottabler)
library(openxlsx)
library(purrr)
library(overlap)
library(data.table)
library(circular)
```

```{r}
bob_coy <- read.csv("bob_coy_overlap_noJan2021.csv")
str(bob_coy)

bob_coy <- bob_coy |>
  filter(Station != "2_AEP")

str(bob_coy)
```

## ANNUAL ACTIVITY

Investigate activity of bobcats and coyotes separately and their overlap
```{r}
bobcat <- bob_coy[bob_coy$Species == "Bobcat",]$radtime
coyote <- bob_coy[bob_coy$Species == "Coyote",]$radtime

densityPlot(bobcat, rug = TRUE)
densityPlot(coyote, rug = TRUE)

#tiff("annual overlap.tif", width = 7, height = 5, units = "in", res = 200)
overlapPlot(bobcat, coyote, main= "Annual Activity")
legend('top', c("Bobcat", "Coyote"), lty=c(1,2), col=c(1, "blue3"), bty='n',cex = 1)
#dev.off()

```

Estimate overlap coefficient
```{r}
bob_coy_est <- overlapEst(bobcat, coyote, type = ("Dhat4"))
print(bob_coy_est)

watson.wheeler.test(list(bobcat, coyote))
```

Get confidence intervals for overlap coefficient via bootstrapping
```{r}
bs <- bootstrap(bobcat, coyote, 10000, type = "Dhat4")
mean(bs)
```

```{r}
hist(bs)
abline(v=bob_coy_est, col='red', lwd=2) # our original estimate
abline(v=mean(bs), col='blue', lwd=2, lty=3) # the estimate based on the 10000 simulated values
```

Get CIs for the bootstrapping estimate
```{r}
(CInorm <- bootCI(bob_coy_est, bs)['norm0', ])

hist(bs, main="Bootstrap Estimation of Overlap Coefficient", cex.main=2)
abline(v=bob_coy_est, col='red', lwd=2) # our original estimate
abline(v=mean(bs), col='blue', lwd=2, lty=3) # the estimate based on the 10000 simulated values
abline(v=CInorm, col='deeppink', lwd=3, lty=3) # the confidence intervals
```

## SEASONAL ACTIVITY

Split data between Spring (Feb - May; reprod and pup/kit birth), Summer (May-Sep; kit and pup rearing), and Fall-Winter (Oct-Jan; it and pup dispersal, cold season etc)
```{r}

bob_coy_Spring <- bob_coy |>
  filter(Month %in% c(2:5))  
str(bob_coy_Spring)

# count number of bobcat and coyote detections
bob_coy_Spring %>%
  group_by(Species) %>%
  summarize(count = n())

bob_coy_Summer <- bob_coy |>
  filter(Month %in% c(6:9))  
str(bob_coy_Summer)

# count number of bobcat and  coyote detections
bob_coy_Summer %>%
  group_by(Species) %>%
  summarize(count = n())


bob_coy_FallWinter <- bob_coy |>
  filter(Month %in% c(1,10:12))
str(bob_coy_FallWinter)

# count number of bobcat and  coyote detections
bob_coy_FallWinter %>%
  group_by(Species) %>%
  summarize(count = n())

```

Bobcat and Coyote seasonal
```{r}
bobcat_Spring <- bob_coy_Spring[bob_coy_Spring$Species == "Bobcat",]$radtime
bobcat_Summer <- bob_coy_Summer[bob_coy_Summer$Species == "Bobcat",]$radtime
bobcat_FallWinter <- bob_coy_FallWinter[bob_coy_FallWinter$Species == "Bobcat",]$radtime

coyote_Spring <- bob_coy_Spring[bob_coy_Spring$Species == "Coyote",]$radtime
coyote_Summer <- bob_coy_Summer[bob_coy_Summer$Species == "Coyote",]$radtime
coyote_FallWinter <- bob_coy_FallWinter[bob_coy_FallWinter$Species == "Coyote",]$radtime
```

Bobcat seasonal activity
```{r}
densityPlot(bobcat_Spring, rug = TRUE)
```
```{r}
densityPlot(bobcat_Summer, rug = TRUE)
```
```{r}
densityPlot(bobcat_FallWinter, rug = TRUE)
```

Coyote seasonal activity
```{r}
densityPlot(coyote_Spring, rug = TRUE)
```
```{r}
densityPlot(coyote_Summer, rug = TRUE)
```
```{r}
densityPlot(coyote_FallWinter, rug = TRUE)
```

Bobcat and coyote seasonal overlap

```{r}
overlapPlot(bobcat_Spring, coyote_Spring, main="Spring")
legend('top', c("Bobcat", "Coyote"), lty=c(1,2), col=c(1, "blue3"), bty='n',cex = 1)
```
Estimate overlap coefficient
```{r}
bob_coy_est_Spring <- overlapEst(bobcat_Spring, coyote_Spring, type = ("Dhat4"))
print(bob_coy_est_Spring)

watson.wheeler.test(list(bobcat_Spring, coyote_Spring))
```

```{r}
overlapPlot(bobcat_Summer, coyote_Summer, main="Summer")
legend('top', c("Bobcat", "Coyote"), lty=c(1,2), col=c(1,"blue3"), bty='n',cex = 1)
#mtext(expression(paste(Delta[1] == 0.74, "*", ", ", Delta[4] == 0.73, "*")), side = 3, line = 0.5, cex = 0.8, adj = 0.5)
```
Estimate overlap coefficient
```{r}
bob_coy_est_Summer <- overlapEst(bobcat_Summer, coyote_Summer, type = ("Dhat1"))
print(bob_coy_est_Summer)

watson.wheeler.test(list(bobcat_Summer, coyote_Summer))
```

```{r}
overlapPlot(bobcat_FallWinter, coyote_FallWinter, main="Fall/Winter")
legend('top', c("Bobcat", "Coyote"), lty=c(1,2), col=c(1,"blue3"), bty='n',cex = 1)
#mtext(expression(paste(Delta[1] == 0.86, ", ", Delta[4] == 0.86)), side = 3, line = 0.5, cex = 0.8, adj = 0.5)
```
Estimate overlap coefficient
```{r}
bob_coy_est_FallWinter <- overlapEst(bobcat_FallWinter, coyote_FallWinter, type = ("Dhat1"))
print(bob_coy_est_FallWinter)

watson.wheeler.test(list(bobcat_FallWinter, coyote_FallWinter))
```

## Let's investigate overlap by region (AEP vs ZAL vs VIN)

AEP LANDS
```{r}
bob_coy1 <- bob_coy %>%
    filter(str_detect(Station, "AEP"))

str(bob_coy1)

# count number of bobcat and coyote detections
bob_coy1 %>%
  group_by(Species) %>%
  summarize(count = n())

# count number of bobcat and coyote detections by camera
bob_coy1_Station <- bob_coy1 %>%
  group_by(Station, Species) %>%
  summarize(count = n()) |>
  rename("Detections" = "count") |>
  as.data.frame() 

ggplot(data = bob_coy1_Station, aes(x = reorder(Station, Detections), y = Detections, fill = Species)) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.75)  +
  geom_text(aes(label = Detections), fontface = "bold", vjust = -0.5,  # Move labels above bars
            position = position_dodge(.9), size = 4) +
  labs(x = "\n Station", y = "Camera Events \n", title = "\n AEP ReCreation Lands \n") +
  scale_fill_manual(values = c("Bobcat" = "black", "Coyote" = "grey")) +  # Adjust colors
  ylim(0, 16) +  # Set the y-axis limit from 0 to 16
  theme(plot.title = element_text(hjust = 0.5), 
        axis.title.x = element_text(face="bold", size = 12),
        axis.title.y = element_text(face="bold", size = 12),
        legend.title = element_text(face="bold", size = 10)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Overlap graph
```{r}
bobcat1 <- bob_coy1[bob_coy1$Species == "Bobcat",]$radtime
coyote1 <- bob_coy1[bob_coy1$Species == "Coyote",]$radtime

overlapPlot(bobcat1, coyote1, main="AEP ReCreation Lands")
legend('top', c("Bobcat", "Coyote"), lty=c(1,2), col=c(1,"blue3"), bty='n',cex = 1)
#mtext(expression(paste(Delta[4] == 0.82, "*")), side = 3, line = 0.5, cex = 0.8, adj = 0.5)
```

Estimate overlap coefficient AEP Lands
```{r}
bob_coy_est1 <- overlapEst(bobcat1, coyote1, type = ("Dhat4"))
print(bob_coy_est1)

watson.wheeler.test(list(bobcat1, coyote1))
```

ZALESKI STATE FOREST
```{r}
bob_coy2 <- bob_coy %>%
  filter(str_detect(Station, "ZAL"))

str(bob_coy2)

# count number of bobcat and coyote detections
bob_coy2 %>%
  group_by(Species) %>%
  summarize(count = n())


# count number of bobcat and coyote detections by camera
bob_coy2_Station <- bob_coy2 %>%
  group_by(Station, Species) %>%
  summarize(count = n()) |>
  rename("Detections" = "count") |>
  as.data.frame() 

# let's check out number of detections of bobcats and coyotes by camera to evaluate potential spatial avoidance 
ggplot(data = bob_coy2_Station, aes(x = reorder(Station, Detections), y = Detections, fill = Species)) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.75)  +
 # ylim(0,30) +
  geom_text(aes(label = Detections), fontface = "bold", vjust = -0.5,
             position = position_dodge(.9), size = 4) +
  labs(x = "\n Station", y = "Camera Events \n", title = "\n Zaleski State Forest \n") +
   scale_fill_manual(values = c("Bobcat" = "black", "Coyote" = "darkgrey")) +  # Adjust colors
   ylim(0, 53) + 
  theme(plot.title = element_text(hjust = 0.5), 
        axis.title.x = element_text(face="bold", size = 12),
        axis.title.y = element_text(face="bold", size = 12),
        legend.title = element_text(face="bold", size = 10)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Overlap plot
```{r}
bobcat2 <- bob_coy2[bob_coy2$Species == "Bobcat",]$radtime
coyote2 <- bob_coy2[bob_coy2$Species == "Coyote",]$radtime

overlapPlot(bobcat2, coyote2, main="Zaleski State Forest")
legend('top', c("Bobcat", "Coyote"), lty=c(1,2), col=c(1,"blue3"), bty='n',cex = 1)
#mtext(expression(paste(Delta[4] == 0.77, "*")), side = 3, line = 0.5, cex = 0.8, adj = 0.5)
```

Estimate overlap coefficient Zaleski
```{r}
bob_coy_est2 <- overlapEst(bobcat2, coyote2, type = ("Dhat4"))
print(bob_coy_est2)

watson.wheeler.test(list(bobcat2, coyote2))
```

VINTON-FURNACE FOREST
```{r}
bob_coy3 <- bob_coy %>%
  filter(str_detect(Station, "VIN"))

str(bob_coy3)

# count number of bobcat and coyote detections
bob_coy3 %>%
  group_by(Species) %>%
  summarize(count = n())

# count number of bobcat and coyote detections by camera
bob_coy3_Station <- bob_coy3 %>%
  group_by(Station, Species) %>%
  summarize(count = n()) |>
  rename("Detections" = "count") |>
  as.data.frame() 

# let's check out number of detections of bobcats and coyotes by camera to evaluate potential spatial avoidance 
ggplot(data = bob_coy3_Station, aes(x = reorder(Station, Detections), y = Detections, fill = Species)) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.75)  +
 # ylim(0,30) +
  geom_text(aes(label = Detections), fontface = "bold", vjust = -0.5,
             position = position_dodge(.9), size = 4) +
  labs(x = "\n Station", y = "Camera Events \n", title = "\n Vinton-Furnace State Forest \n") +
  scale_fill_manual(values = c("Bobcat" = "black", "Coyote" = "darkgrey")) +  # Adjust colors
   ylim(0, 34) + 
  theme(plot.title = element_text(hjust = 0.5), 
        axis.title.x = element_text(face="bold", size = 12),
        axis.title.y = element_text(face="bold", size = 12),
        legend.title = element_text(face="bold", size = 10)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Overlap plot
```{r}
bobcat3 <- bob_coy3[bob_coy3$Species == "Bobcat",]$radtime
coyote3 <- bob_coy3[bob_coy3$Species == "Coyote",]$radtime

overlapPlot(bobcat3, coyote3, main="Vinton-Furnace State Forest")
legend('top', c("Bobcat", "Coyote"), lty=c(1,2), col=c(1,"blue3"), bty='n',cex = 1)
#mtext(expression(paste(Delta[1] == 0.88, ", ", Delta[4] == 0.88)), side = 3, line = 0.5, cex = 0.8, adj = 0.5)
```

Estimate overlap coefficient Vinton
```{r}
bob_coy_est3 <- overlapEst(bobcat3, coyote3, type = ("Dhat1"))
print(bob_coy_est3)

watson.wheeler.test(list(bobcat3, coyote3))
```
